ms_5

### 2020-04-25 ms_5.Rmd
models and new data

library(dplyr)
library(janitor)
library(tidyverse)
library(ggthemes)
library(maps)
library(leaflet)
library(gganimate)
```

@@

@@

ggplot(data = treatment_locations_map_per_cap,
            mapping = aes(x = long, y = lat.y,
                          fill = count_per_pop, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
       color = "Number of Treatment Centers in State Per Capita")
```

@@

>>>>

library(dplyr)
library(janitor)
library(tidyverse)
library(ggthemes)
library(maps)
library(leaflet)
library(gganimate)
library(broom)
```

@@

# not in final

@@

```{r}
# in final

ggplot(data = treatment_locations_map_per_cap,
            mapping = aes(x = long, y = lat.y,
                          fill = count_per_pop, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
       color = "Number of Treatment Centers in State Per Capita")

@@

# models generated from data from kff.org

# ageopioid

ageopioidmodel <- read_csv("U.S. age vs. opioid deaths.csv", col_types = "dcc") %>%
  clean_names() %>%
  na.omit() %>%
  filter(age_range != "Total") %>%
  mutate(age_range = as.factor(age_range)) %>%
    mutate(number_of_deaths = as.numeric(number_of_deaths))

ggplot(ageopioidmodel, aes(year, number_of_deaths, color = age_range)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Age Range")

#raceopioid

raceopioidmodel <- read_csv("U.S. race vs. opioid deaths.csv", col_types = "dcc") %>%
  clean_names() %>%
  na.omit() %>%
  mutate(race = as.factor(race)) %>%
  mutate(opioid_deaths = as.numeric(opioid_deaths))

ggplot(raceopioidmodel, aes(year, opioid_deaths, color = race)) +
  geom_point() +
  scale_color_viridis_d() +
  theme_classic() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Year", y = "Number of Opioid Deaths in America", color = "Race")
```

# filtering data by race/age
ages0_24 <- ageopioidmodel %>%
  filter(age_range == "0-24")

ages0_24 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages25_34 <- ageopioidmodel %>%
  filter(age_range == "25-34")

ages25_34 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages35_44 <- ageopioidmodel %>%
  filter(age_range == "35-44")

ages35_44 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages45_54 <- ageopioidmodel %>%
  filter(age_range == "45-54")

ages45_54 %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

ages55_up <- ageopioidmodel %>%
  filter(age_range == "55+")

ages55_up %>%
    lm(number_of_deaths ~ year, data = .) %>% 
    tidy(conf.int = TRUE) %>% 
    select(term, estimate, conf.low, conf.high)

# averageageopioidmodel <- ageopioidmodel %>%
  # drop_na() %>%
  # group_by(age_range) %>%
  # summarize(mean_deaths = mean(number_of_deaths))

# averageageopioid_model <- lm(mean_deaths ~ age_range, data = averageageopioidmodel)
# averageageopioid_model %>%
  # tidy(conf.int = TRUE) %>%
  # select(term, estimate, conf.low, conf.high)
```
